---
title: "Practical Rates and Rates Ratio"
author: "Ken Mwai"
theme: Szeged
output:
  slidy_presentation: default
fontsize: 14pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, tidy = TRUE, 
      tidy.opts=list(blank=FALSE, width.cutoff=60),
      width = 150)
library(foreign)
library(epiDisplay)
library(epitools)
library(epiR)
library(fmsb)
```

## 1.0 Load packages

```{r, tidy=T, eval=FALSE}
library(foreign)
library(epiDisplay)
library(epitools)
library(epiR)
library(fmsb)
```

## 2.0 Load the child deaths data
```{r, tidy=T}
childdeaths <- read.table("child.deaths.csv", header=TRUE, sep=",")
str(childdeaths)
```


## 2.1 label the values of the variable "status"

```{r, tidy=T}
childdeaths$status<-as.numeric(childdeaths$status)
childdeaths$status <- factor(childdeaths$status,
                             levels = c(1,2),
                             labels = c("alive", "dead"))
table(childdeaths$status)
```


## 3.0 generate person-years


```{r, tidy=T}
childdeaths$date_exit <- as.Date(childdeaths$date_exit, "%m/%d/%y")
childdeaths$date_birth <-
  as.Date(childdeaths$date_birth, "%m/%d/%y")

person_yrs <- difftime(childdeaths$date_exit, childdeaths$date_birth)
person_yrs <- (as.numeric(person_yrs)) / 365.25
childdeaths <- data.frame(childdeaths, person_yrs)

```


***
```{r, tidy=T}
head(childdeaths)
```

## 4.0 Calculate rates

```{r, tidy=T}
total_deaths <- length(childdeaths$status[childdeaths$status == "dead"])
total_deaths
```


```{r, tidy=T}
total_pyrs<-sum(childdeaths$person_yrs)
total_pyrs
```


```{r, tidy=T}
mort_rate<-(as.numeric(total_deaths)/total_pyrs)
mort_rate
```


## 5.0 calculate standard error of the rate

```{r, tidy=T}
s.e.Rate<-sqrt(total_deaths)/total_pyrs*1000 # 
s.e.Rate
```


```{r, tidy=T}
s.e.log.Rate<-1/sqrt(total_deaths)
s.e.log.Rate
```


## 6.0 Compute 95% CI for rate

```{r, tidy=T}
log.rate=log(mort_rate)
lower.bound=exp(log.rate-1.96*s.e.log.Rate)
lower.bound
```


```{r, tidy=T}
upper.bound=exp(log.rate+1.96*s.e.log.Rate)
upper.bound
```


# COMPARING TWO RATES

## 7.0 Rate difference: Males vs. Females

- Calculate number of deaths for each gender: female

```{r, tidy=T}
f.deaths <-
  length(childdeaths$status[childdeaths$status == "dead" &
                              childdeaths$sex == "f"])
f.deaths
```

- Calculate number of deaths for each gender: male 

```{r, tidy=T}
m.deaths <-
  length(childdeaths$status[childdeaths$status == "dead" &
                              childdeaths$sex == "m"])
m.deaths
```


## 7.2  calculate person years for each gender

```{r, tidy=T}
f.total_pyrs <- sum(childdeaths$person_yrs[childdeaths$sex == "f"])
f.total_pyrs
```


```{r, tidy=T}
m.total_pyrs <- sum(childdeaths$person_yrs[childdeaths$sex == "m"])
m.total_pyrs
```


***

```{r, tidy=T}
table(childdeaths$status,childdeaths$sex)
```

## 7.3 Calculate rate difference

```{r, tidy=T}
diff1<-(f.deaths/f.total_pyrs)-(m.deaths/m.total_pyrs)
diff1
```

```{r, tidy=T}
gender_diff <-
  ratedifference(f.deaths, m.deaths, f.total_pyrs, m.total_pyrs, conf.level =0.95)
gender_diff
```

# 8.0 Rate Ratio: Home vs. Hospital delivery

## 8.1 Calculate number of deaths for each gender: male and female

```{r, tidy=T}
home_deaths <-
  length(childdeaths$status[childdeaths$status == "dead" &
                              childdeaths$locn_birth == "Home"])
hosp_deaths <-
  length(childdeaths$status[childdeaths$status == "dead" &
                              childdeaths$locn_birth == "Hospital"])
table(childdeaths$status, childdeaths$locn_birth)
```

## 8.2  calculate person years for both babies delivered in hospital and at home

```{r , tidy=T}
home_pyrs <-
  sum(childdeaths$person_yrs[childdeaths$locn_birth == "Home"])
hosp_pyrs <-
  sum(childdeaths$person_yrs[childdeaths$locn_birth == "Hospital"])
home_vs_hosp <- ratetable(home_deaths, hosp_deaths, home_pyrs, hosp_pyrs)
dimnames(home_vs_hosp) <-
  list(Exposure = c("Home", "Hospital"),
       Outcome = c("Deaths", "PYears"))
home_vs_hosp
```

## 8.3  calculate the rate ratio

```{r , tidy=T}
rateratio.wald(home_vs_hosp,rev="r")
```

## 8.4 Rate Ratio: mother's education (some education Vs. no education)

```{r}
no_educ_deaths<-length(childdeaths$status[childdeaths$status=="dead" & childdeaths$mom_educ=="no education"])
some_educ_deaths<-length(childdeaths$status[childdeaths$status=="dead" & childdeaths$mom_educ=="some education"])
table(childdeaths$status,childdeaths$mom_educ)

```


***

```{r, tidy=T}
no_educ_pyrs<-sum(childdeaths$person_yrs[childdeaths$mom_educ=="no education"])
some_educ_pyrs<-sum(childdeaths$person_yrs[childdeaths$mom_educ=="some education"])
some_educ_Vs_no_educ<-ratetable(no_educ_deaths,some_educ_deaths,no_educ_pyrs,some_educ_pyrs)
dimnames(some_educ_Vs_no_educ)<-list(Exposure=c("no education","some education"), Outcome=c("Deaths","PYears"))
some_educ_Vs_no_educ
```

***


```{r, tidy=T}
rateratio.wald(some_educ_Vs_no_educ,rev="r")
```
