---
title: "Linear Regression Examples"
author: "Ken Mwai"
date: "3/19/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
library(haven)
```

## Data 
1.	We are going to use a dataset from an HIV study. Read the dataset “hivinfection weight” into R We will be using the variables: 
  a.	Weight -  child birth weight
  b.	CD4 – maternal CD4 count at delivery
  c.	Momhb – Maternal Haemoglobin levels at delivery
  d.	ga_weeks – gestational age at delivery/birth
  e.	hiv_status – Child’s HIV status at birth
  f.	sex  - the sex of the child
  
2.	The primary research question is __“Is child birth weight associate with maternal CD4 count at birth?”__

3.	What other questions would be worth asking in this study based on the variables given? 

4.	Let’s define these questions and use these questions to base our practical


## Child birth weight associate with maternal CD4 count

```{r}
hiv_df <- read_dta(file = "hivdata.dta")

## structure of the data
glimpse(hiv_df)
```



