---
title: "Linear Regression Practical Overview"
author: "Ken Mwai"
date: "3/19/2021"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#options(scipen=999)
options(scipen = 999, digits =4) #set to two decimal 
```


```{r}
library(tidyverse)
library(haven)
library(broom)## tidy model reports
```

## Data 
1.	We are going to use a dataset from an HIV study. Read the dataset “hivinfection weight” into R We will be using the variables: 
  a.	weight_kg -  child birth weight
  b.	cd4 – maternal CD4 count at delivery
  c.	momhb – Maternal Haemoglobin levels at delivery
  d.	ga_weeks – gestational age at delivery/birth
  e.	hiv_status – Child’s HIV status at birth
  f.	sex  - the sex of the child
  
2.	The primary research question is __“Is child birth weight associated with child HIV status?”__

- Adjust for any confounders and fit a multi-variable model

3.	The primary research question is __“Is child birth weight associated with maternal CD4 count at birth?”__

- Adjust for any confounders and fit a multi-variable model

## Child birth weight associate with maternal HIV status

```{r}
hiv_df <- read_dta(file = "hivdata.dta")

## structure of the data
glimpse(hiv_df)
```


## Fitting the model

```{r}
lm1 <- lm(weight_kg ~ hiv_status , data = hiv_df)
summary(lm1)
```


## Diagnostic plots for model 1

```{r}
par(mfrow = c(2, 2))
plot(lm1)
```


### Tidy model

```{r}
confint(lm1)
tidy(lm1 ,  conf.int = T )
```

### Extract the AIC

```{r}
glance_lm1 <- glance(lm1)
glance_lm1
```

## Report

On average we observe that the HIV positive individuals have a significantly lower birth weight of 0.18kg (95% C.I =(-0.33,	-0.03), pval=0.01)  compared to the HIV negative individuals. 


### Adjsut for gestation weeks

```{r}
lm2 <- lm(weight_kg ~ hiv_status + ga_weeks , data = hiv_df)
confint(lm2)
tidy(lm2 , conf.int = T )
```

## Diagnostic plots for model 2

```{r}
par(mfrow = c(2, 2))
plot(lm2)
```


```{r}
glance_lm2 <- glance(lm2)
glance_lm2
```


## Interaction predictors

```{r}
##https://cran.r-project.org/web/packages/interactions/vignettes/interactions.html
lm3 <- lm(weight_kg ~ hiv_status + ga_weeks + hiv_status*ga_weeks , data = hiv_df)
tidy(lm3 , conf.int = T )
```


## Diagnostic plots for model 3

```{r}
par(mfrow = c(2, 2))
plot(lm3)
```

## Child birth weight associate with maternal CD4 count

Check if there is a linear association between the outcome and the predictor

```{r}
ggplot(data = hiv_df ,mapping = aes(x = cd4,
                                y = weight_kg )) +
  #Add a point geom
  geom_point() +
  ##Make it straight
  geom_smooth(method = "lm") +
  ## add labels
  labs(x = "Maternal CD4 count", y = "Birthweight (kgs)",
       title = "Birth weight by maternal CD4 count")+
  theme_bw()

```



```{r}
lm4 <- lm(weight_kg ~ cd4  , data = hiv_df)
#confint(lm4)
broom::tidy(lm4 , conf.int = T )
```


- Diagnostic plots model 4 ?? . We create a CD4 count in 100's


```{r}
hiv_df <- hiv_df %>% 
  mutate(cd4_100=cd4/100)
```



```{r}
lm5 <- lm(weight_kg ~ cd4_100 , data = hiv_df)
#confint(lm4)
broom::tidy(lm5 , conf.int = T ) 
```


## Adjust for confounders

```{r}
lm6 <- lm(weight_kg ~ cd4_100 + sex + ga_weeks +momhb , data = hiv_df)
#confint(lm4)

```

- Check how the sample size reduces if you fit the model using a predictor with missing data points. Compare model `lm6` and `lm5`


```{r}
summary(lm6)
```


```{r}
summary(lm5)
```

